---
title: "MCMC"
author: "Yaqiong Yao"
date: "10/26/2018"
output: pdf_document
---

```{r}
require("HI")
require("MCMCpack")
```


We know that :

$$ x| \delta, \mu_1, \mu_2, \sigma_1^2, \sigma_2^2 \sim \delta N(\mu_1, \sigma_1^2) + (1-\delta)N(\mu_2, \sigma_2^2)$$

The prior distributions for five parameters are:

$$\pi(\delta) = 1$$
$$ \mu_1 \sim N(0, 100)$$
$$ \mu_2 \sim N(0, 100)$$
$$\sigma_1^2 \sim IG(0.5, 10)$$
$$\sigma_2^2 \sim IG(0.5, 10)$$

The full conditional distribution :
\begin{align*}
f(\delta, \mu_1, \mu_2, \sigma_1^2, \sigma_2^2|\mathbf{x}) \propto \prod_{i=1}^n \left( \delta\frac{1}{\sqrt{2\pi\sigma_1^2}}exp\left\{ -\frac{(x_i - \mu_1)^2}{2\sigma_1^2} \right\} + (1-\delta)\frac{1}{\sqrt{2\pi\sigma_2^2}}exp\left\{ -\frac{(x_i - \mu_2)^2}{2\sigma_2^2} \right\} \right) \\
\times exp\left\{ -\frac{\mu_1^2}{200} \right\} \times exp\left\{ -\frac{\mu_2^2}{200} \right\} \times (\sigma_1^2)^{-1.5}exp\left\{ -\frac{(\sigma_1^2)^2}{10}\right\}\times (\sigma_2^2)^{-1.5}exp\left\{ -\frac{(\sigma_2^2)^2}{10}\right\}
\end{align*}


Take log on both side to get the loglikelihood function.



```{r}
delta <- 0.7 
n <- 100
set.seed(123)
u <- rbinom(n, prob = delta, size = 1)
x <- rnorm(n, ifelse(u == 1, 7, 10), 0.5)
```



```{r}
mylike <- function(theta, x){
  delta <- theta[1]
  mu1 <- theta[2]; sigma21 <- theta[3]
  mu2 <- theta[4]; sigma22 <- theta[5]
  return(sum(log(delta*dnorm(x, mu1, sqrt(sigma21)) + (1-delta)*dnorm(x, mu2, sqrt(sigma22)))) +
    log(dnorm(mu1, 0, 10)) + log(dnorm(mu2, 0, 10)) + log(dinvgamma(sigma21, shape = 0.5, scale = 10)) +
    log(dinvgamma(sigma22, shape = 0.5, scale = 10)))
}
```

```{r}
mymcmc <- function(niter, thetaInit, x) {
  p <- length(thetaInit)
  thetaCurrent <- thetaInit
  out <- matrix(NA, nrow = niter, ncol = p)
  for (i in 1:niter) {
    for (j in 1:p) {
      logFC <- function(thj) {
        theta <- thetaCurrent
        theta[j] <- thj
        mylike(theta, x)
      }
      
      if(j == 1){
        out[i, j] <- thetaCurrent[j] <-
        HI::arms(thetaCurrent[j], logFC, 
                 function(x) ((x > 0) * (x < 1)), 1)
      } else if(j %in% c(2, 4)){
        out[i, j] <- thetaCurrent[j] <-
        HI::arms(thetaCurrent[j], logFC, 
                 function(x) ((x > -10) * (x < 20)), 1)
      } else if(j %in% c(3, 5)){
        out[i, j] <- thetaCurrent[j] <-
        HI::arms(thetaCurrent[j], logFC, 
                 function(x) ((x > 0) * (x < 10)), 1)
      }
    }
  }
  out
}
niter <- 5000
thetaInit <- c(0.5, 0, 1, 0, 1)
set.seed(1026)
result <- mymcmc(niter, thetaInit, x)

plot(ts(result[,1]), ylab = "delta")
plot(ts(result[,2]), ylab = "mu1")
plot(ts(result[,3]), ylab = "sigma21")
plot(ts(result[,4]), ylab = "mu2")
plot(ts(result[,5]), ylab = "sigma22")

hist(result[1001:5000, 1], main = "delta")
hist(result[1001:5000, 2], main = "mu1")
hist(result[1001:5000, 3], main = "sigma21")
hist(result[1001:5000, 4], main = "mu2")
hist(result[1001:5000, 5], main = "sigma22")
```

I only take last 4000 result in histogram and the result we gets from mcmc is similar to the true value.
